Supplemental Information B: Full Annotated Analysis and Reproducible Code

### 2022-02-01 Supplemental-Material-B-Annotated-Analysis.Rmd
supplement B re-renders

\begin{equation}
(\#eq:workerOutSlopeAttQltNeedEq)
(\#eq:WorkerRandomSlopes)
  \begin{split}
    \textrm{Level 1:} &
      \begin{aligned}[t]
        \ Attitude_{ti} =  &\ \beta_{0i} + \beta_{1i}KeyNeedFulfill_{ti} + \\
                           &\ \beta_{2i}InteractionQuality_{ti} + e_{ti}
      \end{aligned} \\
    \textrm{Level 2:} &\ \beta_{0i} = \gamma_{00} + u_{0i} \\
                      &\ \beta_{1i} = \gamma_{10} + u_{1i} \\
                      &\ \beta_{2i} = \gamma_{20} + u_{2i}
  \end{split} 
\end{equation}

@@

# Get summary with p-values (Satterthwaite's method)
summ(
  lmer(AttitudesDutch ~ KeyNeedFullfillment_cwc + (1 | PID), 
       data = studentOutWithinBetween),
  confint = TRUE,
  digits = 3,
  center = TRUE
)

@@

# Get summary with p-values (Satterthwaite's method)
summ(
  lmer(
    AttitudesDutch ~ KeyNeedFullfillment_cwc + quality_overall_cwc + (1 | PID),
    data = studentOutWithinBetween
  ),
  confint = TRUE,
  digits = 3,
  center = FALSE
)

@@

\begin{equation}
(\#eq:workerOutSlopeAttQltNeedEq)
(\#eq:WorkerRandomSlopes)
  \begin{split}
    \textrm{Level 1:} &
      \begin{aligned}[t]
        \ Attitude_{ti} =  &\ \beta_{0i} + \beta_{1i}KeyNeedFulfill_{ti} + \\
                           &\ \beta_{2i}InteractionQuality_{ti} + e_{ti}
      \end{aligned} \\
    \textrm{Level 2:} &\ \beta_{0i} = \gamma_{00} + u_{0i} \\
                      &\ \beta_{1i} = \gamma_{10} + u_{1i} \\
                      &\ \beta_{2i} = \gamma_{20} + u_{2i}
  \end{split} 
\end{equation}

@@

@@

@@

@@

@@

@@

@@

@@

We sequentially tested whether the fulfillment of Allport's Contact Conditions was (1) related to more positive outgroup attitudes, (2) higher perceived interaction quality, and (3) whether the variance explained by Allport's Conditions is assumed by the perceived interaction quality if considered jointly. We find that in the multilevel models, the fulfillment of Allport's Conditions during outgroup contacts was associated with more positive outgroup attitudes (random slopes model; `r MlCoeffHtml(lmeMdl = mdlMedicalOut$Att$lmeSlopesAllport, lmerCI = mdlMedicalOut$Att$lmerSlopesAllportCI, varName = "AllportsCondition_cwc")`) and also predicted higher perceived interaction quality (random slopes model; `r MlCoeffHtml(lmeMdl = mdlMedicalOut$Qlt$lmeSlopesAllport, lmerCI = mdlMedicalOut$Qlt$lmerSlopesAllportCI, varName = "AllportsCondition_cwc")`). However, if we consider the influences of Allport's Conditions and interaction quality on outgroup attitudes jointly, we find that perceived interaction quality is a substantially stronger predictor (random slopes mode; `r MlCoeffHtml(lmeMdl = mdlMedicalOut$Att$lmeSlopesAllportQlt, lmerCI = mdlMedicalOut$Att$lmerSlopesAllportQltCI, varName = "qualityOverall_cwc")`) and the unique variance explained by Allport's Conditions was less than half of its original effect size (`r MlCoeffHtml(lmeMdl = mdlMedicalOut$Att$lmeSlopesAllportQlt, lmerCI = mdlMedicalOut$Att$lmerSlopesAllportQltCI, varName = "AllportsCondition_cwc")`). These results indicate that in this dataset the the fulfillment of Allport's conditions had a significant influence on outgroup attitudes and this effect is likely in parts explained by its effect through perceived interaction quality.

@@

@@

@@

@@

@@

@@

mdlMedicalOut$Att$lmerInterceptCoreQltCI <- 
  confint(method = "Wald", mdlMedicalOut$Att$lmerInterceptCoreQlt)

@@

@@

We again sequentially tested whether the fulfillment of the core need during an interaction was (1) related to more positive outgroup attitudes, (2) higher perceived interaction quality, and (3) whether the variance explained by the core need is assumed by the perceived interaction quality if considered jointly. We find that in the multilevel models, the fulfillment of core need fulfillment during outgroup contacts was associated with more positive outgroup attitudes (random slopes model; `r MlCoeffHtml(lmeMdl = mdlMedicalOut$Att$lmeSlopesCore, lmerCI = mdlMedicalOut$Att$lmerSlopesCoreCI, varName = "KeyNeedFulfillment_cwc")`) and also predicted higher perceived interaction quality (random slopes model; `r MlCoeffHtml(lmeMdl = mdlMedicalOut$Qlt$lmeSlopesCore, lmerCI = mdlMedicalOut$Qlt$lmerSlopesCoreCI, varName = "KeyNeedFulfillment_cwc")`). However, if we consider the influences of core need fulfillment and interaction quality on outgroup attitudes jointly, we find that perceived interaction quality is a substantially stronger predictor (random slopes mode; `r MlCoeffHtml(lmeMdl = mdlMedicalOut$Att$lmeSlopesCoreQlt, lmerCI = mdlMedicalOut$Att$lmerSlopesCoreQltCI, varName = "qualityOverall_cwc")`) and the unique variance explained by core need fulfillment was roughly half of its original effect size (`r MlCoeffHtml(lmeMdl = mdlMedicalOut$Att$lmeSlopesCoreQlt, lmerCI = mdlMedicalOut$Att$lmerSlopesCoreQltCI, varName = "KeyNeedFulfillment_cwc")`). These results indicate that in this dataset the the fulfillment of core situational needs had a significant influence on outgroup attitudes and this effect is likely in parts explained by its effect through perceived interaction quality.

@@

@@

@@

@@

mdlMedicalOut$Att$lmerSlopesCoreAllportCI <- 
  confint(method = "Wald", mdlMedicalOut$Att$lmerSlopesCoreAllport)

@@

To compare the models using either Allport's conditions or the core need fulfillment to predict outgroup attitudes, we first assess relative model performance indices (i.e., Akaike information criterion, and Bayesian information criterion), and then consider the two predictors in a joint model to see whether the two approaches predict the same variance in outgroup attitudes. When comparing the model selection indices we find that the fulfillment of the situation core need, indeed performs slightly better than the model using Allport's conditions ($AIC_{CoreNeed}$ `r AIC(mdlMedicalOut$Att$lmerSlopesCore)` < `r AIC(mdlMedicalOut$Att$lmerSlopesAllport)` $AIC_{Allport}$, and $BIC_{CoreNeed}$ `r BIC(mdlMedicalOut$Att$lmerSlopesCore)` < `r BIC(mdlMedicalOut$Att$lmerSlopesAllport)` $BIC_{Allport}$). Additionally, when considering the predictors jointly, we find that both significantly predict outgroup attitudes with similar sized regression parameters (random slopes model; Allport's Conditions: `r MlCoeffHtml(lmeMdl = mdlMedicalOut$Att$lmeSlopesCoreAllport, lmerCI = mdlMedicalOut$Att$lmerSlopesCoreAllportCI, varName = "AllportsCondition_cwc")`, Core Need: `r MlCoeffHtml(lmeMdl = mdlMedicalOut$Att$lmeSlopesCoreAllport, lmerCI = mdlMedicalOut$Att$lmerSlopesCoreAllportCI, varName = "KeyNeedFulfillment_cwc")`). This indicates that, although both Allport's conditions and the core need fulfillment seem to (in part) work through perceived interaction quality, they explain different aspects of the variance in outgroup attitudes and do not constitute one another.

@@

```{r medicalModelAttNullInt}
# Create and save Model
mdlMedical$lmerAttNullInt <-
  lme4::lmer(AttitudesDutch ~ 1 + (1 | PID),
             data = dtMedical$full %>%
               filter(complete.cases(
                 OutgroupInteraction
               ))) # use optim if it does not converge

@@

# Get summary with p-values (Satterthwaite's method)
# summary(mdlMedical$lmerAttNull) #or with the lme function
summ(mdlMedical$lmerAttNullInt, digits = 3)
```

```{r medicalModelInterceptAttCoreInt}
# Create and save Model
mdlMedical$lmeInterceptAttCoreInt <-
  lme(
    AttitudesDutch ~ KeyNeedFulfillment_cwc * OutgroupInteraction,
    random =  ~ 1 | PID,
    na.action = na.omit,
    data = dtMedical$full
  )

@@

mdlMedical$lmerInterceptAttCoreIntCI <- 
  confint(method = "Wald", mdlMedical$lmerInterceptAttCoreInt)

@@

@@

We again checked for alternative models to the core need fulfillment. First, when considering generalized situational core need fulfillment together with whether an intergroup contact took place, we find that there is no significant main effect of core need fulfillment (random slopes model; `r MlCoeffHtml(lmeMdl = mdlMedical$lmeSlopesAttCoreInt, lmerCI = mdlMedical$lmerSlopesAttCoreIntCI, varName = "KeyNeedFulfillment_cwc")`) but a stronger interaction effect of core need fulfillment and outgroup contact (`r MlCoeffHtml(lmeMdl = mdlMedical$lmeSlopesAttCoreInt, lmerCI = mdlMedical$lmerSlopesAttCoreIntCI, varName = "KeyNeedFulfillment_cwc:OutgroupInteractionYes")`). Together with a significant main effect of having an outgroup contact (`r MlCoeffHtml(lmeMdl = mdlMedical$lmeSlopesAttCoreInt, lmerCI = mdlMedical$lmerSlopesAttCoreIntCI, varName = "OutgroupInteractionYes")`), this indicates that it is not key need fulfillment in general --- but key need fulfillment during an outgroup contact that predicts more positive outgroup attitudes. This finding is consistent with the results of the previous studies.

>>>>

\begin{equation}
(\#eq:workerOutSlopeAttQltNeedEq)
  \begin{split}
    \textrm{Level 1:} &
      \begin{aligned}[t]
        \ Attitude_{ti} =  &\ \beta_{0i} + \beta_{1i}KeyNeedFulfill_{ti} + \\
                           &\ \beta_{2i}InteractionQuality_{ti} + e_{ti}
      \end{aligned} \\
    \textrm{Level 2:} &\ \beta_{0i} = \gamma_{00} + u_{0i} \\
                      &\ \beta_{1i} = \gamma_{10} + u_{1i} \\
                      &\ \beta_{2i} = \gamma_{20} + u_{2i}
  \end{split} 
\end{equation}

@@

# Get summary with p-values (Satterthwaite's method)
summ(
  mdlStudentOut$Att$lmerInterceptCore <- 
    lmer(AttitudesDutch ~ KeyNeedFullfillment_cwc + (1 | PID), 
       data = studentOutWithinBetween),
  confint = TRUE,
  digits = 3,
  center = TRUE
)

@@

# Get summary with p-values (Satterthwaite's method)
summ(
  mdlStudentOut$Att$lmerInterceptCoreQlt <- lmer(
    AttitudesDutch ~ KeyNeedFullfillment_cwc + quality_overall_cwc + (1 | PID),
    data = studentOutWithinBetween
  ),
  confint = TRUE,
  digits = 3,
  center = FALSE
)

@@

\begin{equation}
(\#eq:studentOutSlopeAttQltNeedEq)
  \begin{split}
    \textrm{Level 1:} &
      \begin{aligned}[t]
        \ Attitude_{ti} =  &\ \beta_{0i} + \beta_{1i}KeyNeedFulfill_{ti} + \\
                           &\ \beta_{2i}InteractionQuality_{ti} + e_{ti}
      \end{aligned} \\
    \textrm{Level 2:} &\ \beta_{0i} = \gamma_{00} + u_{0i} \\
                      &\ \beta_{1i} = \gamma_{10} + u_{1i} \\
                      &\ \beta_{2i} = \gamma_{20} + u_{2i}
  \end{split} 
\end{equation}

@@

We then add the interaction-specific measure of how much Allport's conditions were fulfilled to the multilevel random intercept model.

\begin{equation}
(\#eq:medicalOutInterceptAttAllportEq)
  \begin{split}
    \textrm{Level 1:} &
      \begin{aligned}[t]
        \ Attitude_{ti} =  &\ \beta_{0i} + \beta_{1i}Allport_{ti} + e_{ti}
      \end{aligned} \\
    \textrm{Level 2:} &\ \beta_{0i} = \gamma_{00} + u_{0i} \\
                      &\ \beta_{1i} = \gamma_{10}
  \end{split} 
\end{equation}

@@

We find that the the model with the added predictor indeed explains more variance in outgroup attitudes than the unconditional model. Looking at the individual coefficients, we find that the the fulfillment of Allport's conditions relates significantly to outgroup attitudes.

In a next step, we check whether further letting the effect of Allport's conditions vary between participants explains additional variance in outgroup attitudes (i.e., random slope).

\begin{equation}
(\#eq:medicalSlopeAttAllportEq)
  \begin{split}
    \textrm{Level 1:} &
      \begin{aligned}[t]
        \ Attitude_{ti} =  &\ \beta_{0i} + \beta_{1i}Allport_{ti} + e_{ti}
      \end{aligned} \\
    \textrm{Level 2:} &\ \beta_{0i} = \gamma_{00} + u_{0i} \\
                      &\ \beta_{1i} = \gamma_{10} + u_{1i}
  \end{split} 
\end{equation}

@@

We find that adding the random slopes does add significantly beyond the random intercept model. We also find that Allport's conditions remains a strong predictor (even when letting the influence vary between participants).

### Allport and Interaction Quality

Based on the assertion that the relationship between Allport's conditions and outgroup attitudes is explained by a higher perceived interaction, Allport's conditions should also significantly predict perceived interaction quality.

\begin{equation}
(\#eq:medicalOutQltEmptyModel)
  \begin{split}
    \textrm{Level 1:} &
      \begin{aligned}[t]
        \ InteractionQuality_{ti} =  &\ \beta_{0i} +  e_{ti}
      \end{aligned} \\
    \textrm{Level 2:} &\ \beta_{0i} = \gamma_{00} + u_{0i}
  \end{split} 
\end{equation}

@@

We then add the fulfillment of Allport's conditions to the multilevel random intercept model.

\begin{equation}
(\#eq:medicalOutInterceptQltAllportEq)
  \begin{split}
    \textrm{Level 1:} &
      \begin{aligned}[t]
        \ Attitude_{ti} =  &\ \beta_{0i} + \beta_{1i}Allport_{ti} + e_{ti}
      \end{aligned} \\
    \textrm{Level 2:} &\ \beta_{0i} = \gamma_{00} + u_{0i} \\
                      &\ \beta_{1i} = \gamma_{10}
  \end{split} 
\end{equation}

@@

We find that the the model with the added predictor indeed explains more variance in outgroup attitudes than the unconditional model. Looking at the individual coefficients, we find that the the fulfillment of Allport's conditions relates significantly to perceived interaction quality.

In a next step, we check whether further letting the effect of Allport's conditions vary between participants explains additional variance in outgroup attitudes (i.e., random slope).

\begin{equation}
(\#eq:medicalSlopeQltAllportEq)
  \begin{split}
    \textrm{Level 1:} &
      \begin{aligned}[t]
        \ Attitude_{ti} =  &\ \beta_{0i} + \beta_{1i}Allport_{ti} + e_{ti}
      \end{aligned} \\
    \textrm{Level 2:} &\ \beta_{0i} = \gamma_{00} + u_{0i} \\
                      &\ \beta_{1i} = \gamma_{10} + u_{1i}
  \end{split} 
\end{equation}

@@

We find that adding the random slopes does add significantly beyond the random intercept model. We also find that Allport's conditions remains a strong predictor (even when letting the influence vary between participants).

### Allport, Quality, and Attitudes

In our final main step, we will jointly consider the effect of Allport's conditions and perceived interaction quality on outgroup attitudes. We expect that if the relationship between Allport's conditions and outgroup attitudes is indeed explained by a higher perceived interaction quality, the interaction quality perception should assume the explained variance by the fulfillment of Allport's conditions.

We thus add both Allport's conditions and perceived interaction quality to a random intercept model of outgroup attitudes.

\begin{equation}
(\#eq:medicalOutInterceptAttQltAllportEq)
  \begin{split}
    \textrm{Level 1:} &
      \begin{aligned}[t]
        \ Attitude_{ti} =  &\ \beta_{0i} + \beta_{1i}Allport_{ti} + \\
                           &\ \beta_{2i}InteractionQuality_{ti} + e_{ti}
      \end{aligned} \\
    \textrm{Level 2:} &\ \beta_{0i} = \gamma_{00} + u_{0i} \\
                      &\ \beta_{1i} = \gamma_{10} \\
                      &\ \beta_{2i} = \gamma_{20}
  \end{split} 
\end{equation}

@@

Unsurprisingly, the model with the two predictors adds significantly beyond the empty unconditional model. However, more importantly, looking at the coefficients, we find that the effect of Allport's conditions is indeed is indeed strongly reduced and the variance is explained by the perceived interaction quality. Notably in this sample, the reduction of variance explained is not a much as we saw with core need fulfillment in previous studies.

\begin{equation}
(\#eq:medicalOutSlopeAttQltAllportEq)
  \begin{split}
    \textrm{Level 1:} &
      \begin{aligned}[t]
        \ Attitude_{ti} =  &\ \beta_{0i} + \beta_{1i}Allport_{ti} + \\
                           &\ \beta_{2i}InteractionQuality_{ti} + e_{ti}
      \end{aligned} \\
    \textrm{Level 2:} &\ \beta_{0i} = \gamma_{00} + u_{0i} \\
                      &\ \beta_{1i} = \gamma_{10} + u_{1i} \\
                      &\ \beta_{2i} = \gamma_{20} + u_{2i}
  \end{split} 
\end{equation}

@@

When we consider the influences of Allport's Conditions and interaction quality on outgroup attitudes jointly, we find that perceived interaction quality is a substantially stronger predictor and the unique variance explained by Allport's Conditions was less than half of its original effect size.

@@

The main focus of our proposal is again the assessment of how much positive outgroup attitudes might be explained by whether or not an intergroup interaction fulfilled the person's core situational need. As with the previous two studies we also check to what extend this effect goes through perceived interaction quality.

In a first step we, again, test the relationship between outgroup attitudes and the fulfillment of the core situational need during the interaction.

We jump right in with the random intercept model because we have already assessed the unconditional model when we tested the influence of Allport's conditions on outgroup attitudes. We thus add the core interaction need fulfillment to the multilevel random intercept model.

\begin{equation}
(\#eq:medicalOutInterceptAttNeedEq)
  \begin{split}
    \textrm{Level 1:} &
      \begin{aligned}[t]
        \ Attitude_{ti} =  &\ \beta_{0i} + \beta_{1i}KeyNeedFulfill_{ti} + e_{ti}
      \end{aligned} \\
    \textrm{Level 2:} &\ \beta_{0i} = \gamma_{00} + u_{0i} \\
                      &\ \beta_{1i} = \gamma_{10}
  \end{split} 
\end{equation}

@@

We find that the the model with the added predictor indeed explains more variance in outgroup attitudes than the unconditional model. Looking at the individual coefficients, we find that the situational core need relates significantly to outgroup attitudes.

\begin{equation}
(\#eq:medicalOutSlopeAttNeedEq)
  \begin{split}
    \textrm{Level 1:} &
      \begin{aligned}[t]
        \ Attitude_{ti} =  &\ \beta_{0i} + \beta_{1i}KeyNeedFulfill_{ti} + e_{ti}
      \end{aligned} \\
    \textrm{Level 2:} &\ \beta_{0i} = \gamma_{00} + u_{0i} \\
                      &\ \beta_{1i} = \gamma_{10} + u_{1i}
  \end{split} 
\end{equation}

@@

We again do not need to calculate an unconditional model for the perceived interaction quality outcome variable because we have already done so for the previous assessment of Allport's conditions. So we add the core interaction need fulfillment to the multilevel random intercept model.

\begin{equation}
(\#eq:medicalOutInterceptQltNeedEq)
  \begin{split}
    \textrm{Level 1:} &
      \begin{aligned}[t]
        \ InteractionQuality_{ti} =  &\ \beta_{0i} + \beta_{1i}KeyNeedFulfill_{ti} + e_{ti}
      \end{aligned} \\
    \textrm{Level 2:} &\ \beta_{0i} = \gamma_{00} + u_{0i} \\
                      &\ \beta_{1i} = \gamma_{10}
  \end{split} 
\end{equation}

@@

The predictor again adds a significant amount of explained variances beyond the empty model and looking at the slope coefficient, we find that the situational core need fulfillment relates significantly to perceived interaction quality.

\begin{equation}
(\#eq:medicalOutSlopeQltNeedEq)
  \begin{split}
    \textrm{Level 1:} &
      \begin{aligned}[t]
        \ InteractionQuality_{ti} =  &\ \beta_{0i} + \beta_{1i}KeyNeedFulfill_{ti} + e_{ti}
      \end{aligned} \\
    \textrm{Level 2:} &\ \beta_{0i} = \gamma_{00} + u_{0i} \\
                      &\ \beta_{1i} = \gamma_{10} + u_{1i}
  \end{split} 
\end{equation}

@@

We find that adding the random slopes does add significantly beyond the random intercept model and situation key need fulfillment remains a meaningful predictor of outgroup attitudes.

\begin{equation}
(\#eq:medicalOutInterceptAttQltNeedEq)
  \begin{split}
    \textrm{Level 1:} &
      \begin{aligned}[t]
        \ Attitude_{ti} =  &\ \beta_{0i} + \beta_{1i}KeyNeedFulfill_{ti} + \\
                           &\ \beta_{2i}InteractionQuality_{ti} + e_{ti}
      \end{aligned} \\
    \textrm{Level 2:} &\ \beta_{0i} = \gamma_{00} + u_{0i} \\
                      &\ \beta_{1i} = \gamma_{10} \\
                      &\ \beta_{2i} = \gamma_{20}
  \end{split} 
\end{equation}

@@

# 95%CI
mdlMedicalOut$Att$lmerInterceptCoreQltCI <- 
  confint(method = "Wald", mdlMedicalOut$Att$lmerInterceptCoreQlt)

@@

Unsurprisingly, the model with the two predictors adds significantly beyond the empty unconditional model. However, more importantly, looking at the coefficients, we find that the effect of core need fulfillment indeed is indeed strongly reduced and the variance is explained by the perceived interaction quality. However, similar to the equivalent analysis of Allport's conditions and interaction quality, the effect of need fulfillment remains a significant predictor of outgroup attitudes when controlling for perceived interaction quality.

\begin{equation}
(\#eq:medicalOutSlopeAttQltNeedEq)
  \begin{split}
    \textrm{Level 1:} &
      \begin{aligned}[t]
        \ Attitude_{ti} =  &\ \beta_{0i} + \beta_{1i}KeyNeedFulfill_{ti} + \\
                           &\ \beta_{2i}InteractionQuality_{ti} + e_{ti}
      \end{aligned} \\
    \textrm{Level 2:} &\ \beta_{0i} = \gamma_{00} + u_{0i} \\
                      &\ \beta_{1i} = \gamma_{10} + u_{1i} \\
                      &\ \beta_{2i} = \gamma_{20} + u_{2i}
  \end{split} 
\end{equation}

@@

The random slopes add significant explained variance to the model. When we consider the influences of core need fulfillment and interaction quality on outgroup attitudes jointly, we find that perceived interaction quality is a substantially stronger predictor and the unique variance explained by core need fulfillment was roughly half of its original effect size.

@@

To compare the models using either Allport's conditions or the core need fulfillment to predict outgroup attitudes, we first assess relative model performance indices (i.e., Akaike information criterion, and Bayesian information criterion), and then consider the two predictors in a joint model to see whether the two approaches predict the same variance in outgroup attitudes.

### Model fit parameters

For the model fit parameters we look at the AIC, BIC, pseudo R^2^s, the ICC, RMSE, and Sigma.

@@

When comparing the model selection indices we find that the fulfillment of the situation core need, indeed performs slightly better than the model using Allport's conditions.

### Joint model

We then consider the two predictors in a joint model to see whether the two approaches predict the same variance in outgroup attitudes.

#### random intercept model

We begin by adding both the core need fulfillment and Allport's conditions to a random intercept model of outgroup attitudes.

\begin{equation}
(\#eq:medicalOutInterceptAttNeedAllportEq)
  \begin{split}
    \textrm{Level 1:} &
      \begin{aligned}[t]
        \ Attitude_{ti} =  &\ \beta_{0i} + \beta_{1i}KeyNeedFulfill_{ti} + \\
                           &\ \beta_{2i}Allport_{ti} + e_{ti}
      \end{aligned} \\
    \textrm{Level 2:} &\ \beta_{0i} = \gamma_{00} + u_{0i} \\
                      &\ \beta_{1i} = \gamma_{10} \\
                      &\ \beta_{2i} = \gamma_{20}
  \end{split} 
\end{equation}

@@

For the random intercept model we find that the fulfillment of situational core needs adds significant explained variance beyond a model that only includes Allport's conditions. When looking at the regression parameters in the joint model we additionally find that the fulfillment of core situational needs is ever so slightly a stronger predictor of outgroup attitudes.

\begin{equation}
(\#eq:medicalOutSlopeAttNeedAllportEq)
  \begin{split}
    \textrm{Level 1:} &
      \begin{aligned}[t]
        \ Attitude_{ti} =  &\ \beta_{0i} + \beta_{1i}KeyNeedFulfill_{ti} + \\
                           &\ \beta_{2i}Allport_{ti} + e_{ti}
      \end{aligned} \\
    \textrm{Level 2:} &\ \beta_{0i} = \gamma_{00} + u_{0i} \\
                      &\ \beta_{1i} = \gamma_{10} + u_{1i} \\
                      &\ \beta_{2i} = \gamma_{20} + u_{2i}
  \end{split} 
\end{equation}

@@

# 95%CI
mdlMedicalOut$Att$lmerSlopesCoreAllportCI <- 
  confint(method = "Wald", mdlMedicalOut$Att$lmerSlopesCoreAllport)

@@

Again, the random slopes add significantly to the model. When considering the predictors jointly in the random slopes model, we find that both significantly predict outgroup attitudes with similar sized regression parameters. This indicates that, although both Allport's conditions and the core need fulfillment seem to (in part) work through perceived interaction quality, they explain different aspects of the variance in outgroup attitudes and do not constitute one another.

@@

\begin{equation}
(\#eq:medicalInterceptAttIntNeedEq)
  \begin{split}
    \textrm{Level 1:} &
      \begin{aligned}[t]
        \ Attitude_{ti} =  &\ \beta_{0i} + \beta_{1i}KeyNeedFulfill_{ti} + \\
                           &\ \beta_{2i}OutgroupInteraction_{ti} + \\
                           &\ \beta_{3i}KeyNeedFulfillXOutgroupInteraction_{ti} + e_{ti}
      \end{aligned} \\
    \textrm{Level 2:} &\ \beta_{0i} = \gamma_{00} + u_{0i} \\
                      &\ \beta_{1i} = \gamma_{10} \\
                      &\ \beta_{2i} = \gamma_{20} \\
                      &\ \beta_{3i} = \gamma_{30}
  \end{split} 
\end{equation}

```{r medicalModelInterceptAttCoreInt}
# Create and save empty model for comparison
mdlMedical$lmerAttNullInt <-
  lme4::lmer(AttitudesDutch ~ 1 + (1 | PID),
             data = dtMedical$full %>%
               filter(complete.cases(
                 OutgroupInteraction
               ))) # use optim if it does not converge

@@

# Get summary with p-values (Satterthwaite's method)
#summ(mdlMedical$lmerAttNullInt, digits = 3)

# Create and save Model
mdlMedical$lmeInterceptAttCoreInt <-
  lme(
    AttitudesDutch ~ KeyNeedFulfillment_cwc * OutgroupInteraction,
    random =  ~ 1 | PID,
    na.action = na.omit,
    data = dtMedical$full
  )

@@

# 95%CI
mdlMedical$lmerInterceptAttCoreIntCI <- 
  confint(method = "Wald", mdlMedical$lmerInterceptAttCoreInt)

@@

We find that the model explains significantly more variance than the empty null model. However, more interestingly, looking at the coefficients, we find that, as seen earlier, having an outgroup interaction has a strong effect on outgroup attitudes. Importantly, we find that there is no main effect of key need fulfillment by itself but the effect is qualified by a significant interaction effect of core need fulfillment and outgroup contact. This indicates that it is not simply key need fulfillment in general --- but especially key need fulfillment during an outgroup contact that predicts more positive outgroup attitudes.

\begin{equation}
(\#eq:medicalSlopesAttIntNeedEq)
  \begin{split}
    \textrm{Level 1:} &
      \begin{aligned}[t]
        \ Attitude_{ti} =  &\ \beta_{0i} + \beta_{1i}KeyNeedFulfill_{ti} + \\
                           &\ \beta_{2i}OutgroupInteraction_{ti} + \\
                           &\ \beta_{3i}KeyNeedFulfillXOutgroupInteraction_{ti} + e_{ti}
      \end{aligned} \\
    \textrm{Level 2:} &\ \beta_{0i} = \gamma_{00} + u_{0i} \\
                      &\ \beta_{1i} = \gamma_{10} + u_{1i} \\
                      &\ \beta_{2i} = \gamma_{20} + u_{2i} \\
                      &\ \beta_{3i} = \gamma_{30} + u_{3i}
  \end{split} 
\end{equation}

@@

```{r medicalModelInterceptAttCoreIntPlt}
# visualize interaction
## Without ML structure
ggplot(data = dtMedical$full,
       aes(x = KeyNeedFulfillment,
           y = AttitudesDutch,
           fill = OutgroupInteraction)) +
  #geom_point()+
  geom_smooth(method = 'lm',
              aes(linetype = OutgroupInteraction),
              color = "black") +
  #facet_wrap(~PID, ncol = 6)+
  scale_linetype_manual(values = c("dashed", "solid")) +
  scale_fill_manual(values = c("darkgrey", "black")) +
  #scale_colour_manual(values=c("grey20", "black"), name="Intergroup Contact")+
  scale_y_continuous(
    limits = c(50, 100),
    breaks = seq(50, 100, by = 10),
    position = "left"
  ) +
  scale_x_continuous(limits = c(0, 100), breaks = seq(0, 100, by = 10)) +
  labs(
    title = "Without ML stucture",
    x = "Fulfillment Core Need",
    y = "Outgroup Attitudes",
    fill = "Intergroup Contact",
    linetype = "Intergroup Contact"
  ) +
  theme_Publication() +
  theme(legend.position = "bottom",
        legend.key.size = unit(1, "cm"))

## With ML structure
# create parameters for prediction
datNew = data.frame(
  KeyNeedFulfillment_cwc = rep(seq(
    round_any(min(
      dtMedical$full$KeyNeedFulfillment_cwc %>% na.omit
    ), 5, f = floor), round_any(max(
      dtMedical$full$KeyNeedFulfillment_cwc %>% na.omit
    ), 5, f = ceiling), 5
  ), 2),
  PID = 0
) %>%
  mutate(OutgroupInteraction = rep(c("No", "Yes"), each = nrow(.)/2)) %>%
  select(PID, OutgroupInteraction, KeyNeedFulfillment_cwc)

# Predict values, clean up and calculate SE
PI <-
  merTools::predictInterval(
    merMod = mdlMedical$lmerSlopesAttCoreInt,
    newdata = datNew,
    level = 0.95,
    stat = "mean",
    type = "linear.prediction",
    include.resid.var = F,
    fix.intercept.variance = F
  )
mdlMedical$predInterceptAttCoreIntX <- 
  cbind(datNew, PI)
mdlMedical$predInterceptAttCoreIntX$se <-
  (mdlMedical$predInterceptAttCoreIntX$upr - mdlMedical$predInterceptAttCoreIntX$fit) / 1.96
rm(datNew, PI)
mdlMedical$predInterceptAttCoreIntX$OutgroupInteractionLab <-
  factor(
    x = mdlMedical$predInterceptAttCoreIntX$OutgroupInteraction,
    levels = sort(
      unique(mdlMedical$predInterceptAttCoreIntX$OutgroupInteraction)
    ),
    labels = c("No", "Yes")
  )

# Plot predicted values with SE
ggplot(
  mdlMedical$predInterceptAttCoreIntX,
  aes(x = KeyNeedFulfillment_cwc,
      y = fit,
      fill = OutgroupInteractionLab)
)+
  #geom_point() +
  geom_line(aes(linetype = OutgroupInteractionLab), size = 1) +
  #facet_wrap(~PID, ncol = 6)+
  geom_ribbon(data = mdlMedical$predInterceptAttCoreIntX,
              aes(ymin = fit - se, ymax = fit + se),
              alpha = 0.3) +
  scale_x_continuous(breaks = seq(
    round_any(min(
      studentWithinBetween$KeyNeedFullfillment_cwc
    ), 5, f = floor), round_any(max(
      studentWithinBetween$KeyNeedFullfillment_cwc
    ), 5, f = ceiling), 10
  )) +
  scale_y_continuous(limits = c(50, 100), breaks = seq(50, 100, 10)) +
  scale_linetype_manual(values = c("dashed", "solid")) +
  scale_fill_manual(values = c("darkgrey", "black")) +
  labs(
    x = "Fulfillment Core Need",
    y = "Outgroup Attitude (NL)",
    fill = "Intergroup Contact",
    linetype = "Intergroup Contact",
    title = "Based on Model Predictions"
  ) +
  theme_Publication()
